Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime \thanks{ This work has been supported by the Italian PRIN Project COMMON-WEARS (2020HCWWLP) and the EU/MUR FSE REACT-EU PON R\&I 2014-2020. }

### 2023-05-20 paper-2023-acsos-pulverization.tex
we actually do have space

A framework/runtime configured via declarative files enforces declarativity at the configuration level,
but its flexibility is limited,
as options that have not been accounted for at design time can hardly be injected through configuration files.
%
A (well designed) library can be flexible as the host language can be used to express peculiar configurations
and easier to adopt to users acquainted with the host language;
but its configuration may quickly become imperative and harder to maintain as complexity grows.
%
Finally, a \ac{dsl}~\cite{dsl-book-voelter} can be designed to be as expressive as needed,
but it has high maintenance cost (as the language maintenance costs stack upon those of the library/\ac{api}) and, potentially,
a steep learning curve due to the need of learning a new (custom) language.

@@

In this section,
we exercise the proposed framework through two case studies.
%
In the former,
we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded
to showcase the integration of the framework with established technologies\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-crowd-room}
};
in the latter,
we simulate a city-scale distributed application deployed for a urban event
and show how the dynamic reconfiguration may help achieving balance between energy consumption and \ac{qos}\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}
}.
%
Both experiments have been released with a permissive open-source license and have been archived for future reference on
Zenodo~\cite{https://doi.org/10.5281/zenodo.7933160, https://doi.org/10.5281/zenodo.7948801}.

@@

\subsubsection{Implementation}
We implemented our simulations by interfacing \ourframework{}
with the Alchemist Simulator~\cite{PianiniJOS2013}
through a dedicated platform module.
%
We ran experiments varying the number of participants (hence, smartphones): 100, 300, and 1000
and the \ourframework{} reconfiguration rules:
$\Uparrow$ (behaviour always in the smartphones),
$\Updownarrow^{60}_{20}$,
$\Updownarrow^{80}_{20}$,
$\Updownarrow^{60}_{40}$,
$\Updownarrow^{80}_{40}$,
$\Downarrow$ (behaviour always in the cloud).
%
All the hybrid configurations
(all except $\Uparrow$ and $\Downarrow$)
require runtime reconfiguration to be realised.
%
For each experiment, we run a total of 10 repetitions with a different random seed;
all results are averaged over them.
%
The experiment has been open sourced\footnote{\url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}}
and the code has been archived on Zenodo~\cite{https://doi.org/10.5281/zenodo.7948801}.
%
\begin{figure}
    \includegraphics[width=\linewidth]{figures/simulation-screenshot.png}
    \caption{
        A snapshot of the large-scale simulation.
        Smartphones are represented as coloured circles,
        with the color indicating their battery level
        (green when charged, red when low).
    }
    \label{fig:simulation-screenshot}
\end{figure}
%
\Cref{fig:simulation-screenshot} shows a snapshot of the simulation

\subsubsection{Results}
%
\begin{figure*}
    \includegraphics[width=.32\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}~
    \includegraphics[width=.32\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}~
    \includegraphics[width=.32\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
    \caption{
        Simulation results for 100 (left), 300 (center), and 1000 (right) devices;
        showing all four metrics:
        $P_{cloud}$ ($kW$) in green,
        $P_{devices}$ ($kW$) in blue,
        Distance in yellow, $\$_{cloud}$ in dark purple.
        %
        $P_{cloud}$ and $P_{devices}$ are stacked to also show the overall power consumption.
        %
        Different reconfiguration rules are indicated on the horizontal axis.
        %
        Runtime reconfiguration of the deployment opens to more opportunities to find the trade-off of cost,
        energy use, and \ac{qos} that best fits the current context.
    }
    \label{fig:results}
\end{figure*}
%
Results are summarised in \Cref{fig:results}.
%
$\Downarrow$ has the worst \ac{qos}, but also the lowest cloud cost and overall energy consumption.
%
Opposedly, $\Uparrow$ improves the \ac{qos}, but costs and energy use grow significantly.
%
The ability to reconfigure the deployment at runtime
by moving the behaviour from the cloud to the smartphones and vice versa
depending on the battery level allows to find different trade-offs between these two extremes.

@@

In our approach, the application partitioning into components is manually specified through a \ac{dsl}.
%
Therefore, it can be framed in the context of \emph{component-based software engineering},
 with ~\cite{vale2016component-based-se} providing a comprehensive review.
%
The specification or design of distributed systems
 may leverage
 component models~\cite{DBLP:journals/tse/CrnkovicSVC11},
 architectural description languages~\cite{DBLP:journals/tse/MedvidovicT00},
 service-based compositions~\cite{DBLP:journals/csur/LemosDB16}, 
 or frameworks.
%
Specifically,
 this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation}.
% whereby a large-scale cyber-physical system
% is partitioned into a graph of devices
% where each device is split into five deployable components:
% (i) sensors interface,
% (ii) actuators interface,
% (iii) behaviour,
% (iv) state, and
% (v) communication component.
%
In~\cite{IEEE-IoTJ-pulverization-simulation},
 a methodology on top of the pulverisation model
 is proposed
 for generating and testing deployments through simulation.

@@

In this work,
 we are especially interested in deployments over large-scale cyber-physical systems and the edge-cloud continuum.
%
\emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
 is a \ac{dsl} that allows to express deployment properties of 
applications on multi-scale deployment domains,
i.e., large and heterogeneous infrastructures considered under several viewpoints
(device, geography, network, administrative domains, etc.).
%
In~\cite{DBLP:journals/sosym/SongDFSF22},
  a model-based deployment approach is proposed
  targeting so-called \emph{fleets}, i.e.,
  distributed and heterogeneous devices at the edge characterised by different cyber-physical contexts
(cf. resources, connectivity, etc.).
%
%It is based on the \emph{GeneSIS} modelling language,
%where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.

@@

In the \emph{osmotic computing} approach~\cite{DBLP:journals/computer/VillariFDRJR19},
 microservices (the solvent)
 can migrate across 
 the edge-cloud infrastructure (the solution),
 passing through layer boundaries (the semi-permeable membranes),
 in order to keep a balance in the desired properties (the solute).
%
\emph{ASSL}~\cite{DBLP:conf/birthday/VassevH97}
 supports the specification
 of autonomic elements,
 interaction protocols,
 and \emph{self-management policies} based on fluents (states), events, and actions.
%
%In \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13},
% the language supports the specification of logical \emph{reconfiguration constraints}, which are then solved as a constraint satisfaction problem.
%
The \emph{AWaRE DSL}~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} supports constraint-based self-management,
leveraging managing agents.
%The language enables the specification of the domain model (in terms of components),
%the problem structure model (in terms of constraints),
%the agent architecture model (in terms of agents, roles, and their coordination), and the assignment model (in terms of management problem assignment strategies).
%
Similar, earlier approaches based on the specification of \emph{constraint-sets} include \emph{DELADAS}~\cite{Dearle2004deladas}
and \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13}.
%
\emph{Ctrl-F}~\cite{DBLP:journals/jss/AlvaresRS17}
 is an architectural description language
 with constructs for specifying adaptive behaviour 
 and policies (constraints) for the reconfiguration of system components.
%
These approaches, however, typically suffer from scalability issues.

>>>>

A framework/runtime with configuration files enforces declarativity at the configuration level,
but its flexibility is limited,
as options that have not been accounted for at design time can hardly be injected through configuration files.
%
A (well designed) library can be flexible as the host language can be used to express peculiar configurations
and easier to adopt to users acquainted with the host language;
but its configuration may quickly become imperative and harder to maintain as complexity grows.
%
Finally, a \ac{dsl}~\cite{dsl-book-voelter} can be designed to be as expressive as needed,
but it has high maintenance cost (as the language maintenance costs stack upon those of the library/\ac{api}) and, potentially,
a steep learning curve due to the need of learning a new (custom) language.

@@

In this section,
we exercise the proposed framework through two case studies.
%
In the former,
we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded,
to showcase the integration of the framework with established technologies\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-crowd-room}
};
in the latter,
we simulate a city-scale distributed application deployed for a urban event
and show how the dynamic reconfiguration may help achieving balance between energy consumption and \ac{qos}\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}
}.
%
Both experiments have been released with a permissive open-source license and have been archived for future reference on
Zenodo~\cite{https://doi.org/10.5281/zenodo.7933160, https://doi.org/10.5281/zenodo.7948801}.

@@

\subsubsection{Implementation}
We implemented our simulations by interfacing \ourframework{}
with the Alchemist Simulator~\cite{PianiniJOS2013}
through a dedicated platform module.
%
We ran experiments varying the number of participants (hence, smartphones): 100, 300, and 1000
and the \ourframework{} reconfiguration rules:
$\Uparrow$ (behaviour always in the smartphones),
$\Updownarrow^{60}_{20}$,
$\Updownarrow^{80}_{20}$,
$\Updownarrow^{60}_{40}$,
$\Updownarrow^{80}_{40}$,
$\Downarrow$ (behaviour always in the cloud).
%
All the hybrid configurations
(all except $\Uparrow$ and $\Downarrow$)
require runtime reconfiguration to be realised.
%
For each experiment, we run a total of 10 repetitions with a different random seed;
all results are averaged over them.
%
The experiment has been open sourced
and the code has been archived on Zenodo~\cite{https://doi.org/10.5281/zenodo.7948801}.
%
\begin{figure}
    \includegraphics[width=\linewidth]{figures/simulation-screenshot.png}
    \caption{
        A snapshot of the large-scale simulation.
        Smartphones are represented as coloured circles,
        with the color indicating their battery level
        (green when charged, red when low).
    }
    \label{fig:simulation-screenshot}
\end{figure}
%
\Cref{fig:simulation-screenshot} shows a snapshot of the simulation

\subsubsection{Results}
%
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
    \caption{
        Simulation results for 100 (top), 300 (center), and 1000 (bottom) devices;
        showing all four metrics:
        $P_{cloud}$ ($kW$) in green,
        $P_{devices}$ ($kW$) in blue,
        Distance in yellow, $\$_{cloud}$ in dark purple.
        %
        $P_{cloud}$ and $P_{devices}$ are stacked to also show the overall power consumption.
        %
        Different reconfiguration rules are indicated on the horizontal axis.
        %
        Runtime reconfiguration of the deployment opens to more opportunities to find the trade-off of cost,
        energy use, and \ac{qos} that best fits the current context.
    }
    \label{fig:results}
\end{figure}
%
Results are summarised in \Cref{fig:results}.
%
$\Downarrow$ has the worst \ac{qos}, but also the lowest cloud cost and overall energy consumption.
%
Opposedly, $\Uparrow$ improves the \ac{qos}, but costs and energy use grow significantly.
%
The ability to reconfigure the deployment at runtime
by moving the behaviour from the cloud to the smartphones and vice versa
depending on the battery level allows to find different trade-offs between these two extremes.

@@

In our approach, the application partitioning into components is manually specified through a \ac{dsl}.
%
Therefore, it can be framed in the context of \emph{component-based software engineering},
 with ~\cite{vale2016component-based-se} providing a comprehensive review.
%
The specification or design of distributed systems
 may leverage
 component models~\cite{DBLP:journals/tse/CrnkovicSVC11},
 architectural description languages~\cite{DBLP:journals/tse/MedvidovicT00},
 service-based compositions~\cite{DBLP:journals/csur/LemosDB16}, 
 or frameworks.
%
Specifically,
 this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation}.
 whereby a large-scale cyber-physical system
 is partitioned into a graph of devices
 where each device is split into five deployable components:
 (i) sensors interface,
 (ii) actuators interface,
 (iii) behaviour,
 (iv) state, and
 (v) communication component.
%
In~\cite{IEEE-IoTJ-pulverization-simulation},
 a methodology on top of the pulverisation model
 is proposed
 for generating and testing deployments through simulation.

@@

In this work,
 we are especially interested in deployments over large-scale cyber-physical systems and the edge-cloud continuum.
%
\emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
 is a \ac{dsl} that allows to express deployment properties of 
applications on multi-scale deployment domains,
i.e., large and heterogeneous infrastructures considered under several viewpoints
(device, geography, network, administrative domains, etc.).
%
In~\cite{DBLP:journals/sosym/SongDFSF22},
  a model-based deployment approach is proposed
  targeting so-called \emph{fleets}, i.e.,
  distributed and heterogeneous devices at the edge characterised by different cyber-physical contexts
(cf. resources, connectivity, etc.).
%
It is based on the \emph{GeneSIS} modelling language,
where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.

@@

In the \emph{osmotic computing} approach~\cite{DBLP:journals/computer/VillariFDRJR19},
 microservices (the solvent)
 can migrate across 
 the edge-cloud infrastructure (the solution),
 passing through layer boundaries (the semi-permeable membranes),
 in order to keep a balance in the desired properties (the solute).
%
\emph{ASSL}~\cite{DBLP:conf/birthday/VassevH97}
 supports the specification
 of autonomic elements,
 interaction protocols,
 and \emph{self-management policies} based on fluents (states), events, and actions.
%
%In \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13},
% the language supports the specification of logical \emph{reconfiguration constraints}, which are then solved as a constraint satisfaction problem.
%
The \emph{AWaRE DSL}~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} supports constraint-based self-management,
leveraging managing agents.
The language enables the specification of the domain model (in terms of components),
the problem structure model (in terms of constraints),
the agent architecture model (in terms of agents, roles, and their coordination), and the assignment model (in terms of management problem assignment strategies).
%
Similar, earlier approaches based on the specification of \emph{constraint-sets} include \emph{DELADAS}~\cite{Dearle2004deladas}
and \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13}.
%
\emph{Ctrl-F}~\cite{DBLP:journals/jss/AlvaresRS17}
 is an architectural description language
 with constructs for specifying adaptive behaviour 
 and policies (constraints) for the reconfiguration of system components.
%
These approaches, however, typically suffer from scalability issues.
